On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-08-30 paper.tex
Minor fixes; better Dynkin diagram; proof of Proposition 3.2

\documentclass[oneside, 10pt]{amsart}
\usepackage{amscd, amsmath, amssymb, amsthm, amsfonts, amstext, verbatim, mathtools, xfrac, microtype, nameref, thmtools}
\usepackage[breaklinks=true]{hyperref}
\usepackage[hyperref=true, backend=bibtex, firstinits=true, citestyle=numeric-comp, sortlocale=en_US, url=false, doi=false, eprint=true, maxbibnames=4]{biblatex}
\usepackage[capitalize]{cleveref}
\usepackage[matrix,arrow,curve]{xy}
\usepackage{enumitem}

\addbibresource{paper.bib}
\renewbibmacro*{volume+number+eid}{\ifentrytype{article}{\- \iffieldundef{volume}{}{Vol.~\printfield{volume},}\iffieldundef{number}{}{ No.~\printfield{number},}}}
\renewbibmacro{in:}{\ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}}
\newbibmacro{string+doi}[1]{\iffieldundef{doi}{\iffieldundef{url}{#1}{\href{\thefield{url}}{#1}}}{\href{http://dx.doi.org/\thefield{doi}}{#1}}}
\DeclareFieldFormat[article, inproceedings, inbook, book, thesis]{title}{\usebibmacro{string+doi}{\mkbibquote{#1}}}
\renewcommand*{\bibfont}{\footnotesize}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\T}{T}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\SL}{SL}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\rk}{rk}
\newcommand{\catname}[1]{{\normalfont\textbf{#1}}}

@@

We need one additional technical definition. We call an ordered triple of roots $(\alpha, \beta, \gamma) \in \Phi^3$ an {\it $\rA_3$-triple}
 if these vectors form a basis of a root subsystem $\Psi \subseteq \Phi$ of type $\rA_3$, and moreover if the Dynkin diagram of $\Psi$ with respect to this basis has the form
 $\xymatrix{\circ_\alpha \ar@{-}[r] & \circ_\beta \ar@{-}[r] & \circ_\gamma}.$ %TODO: Draw better diagram :)

@@

We will use the following commutator identities (cf.~\cite[H1]{Re75}):
\begin{align}
% \label{eq:H1i}   [a, bc] = [a, b]\cdot {}^b[a, c] = [a, b] \cdot [b, [a, c]] \cdot [a, c];& \\
 \label{eq:H1ii}  [ab, c] = {}^a[b, c] \cdot [a,c];&\\ %= [a,[b,c]]\cdot [b,c] \cdot [a,c];&\\
 \label{eq:H1iii} [a,c]   = 1    \text{ implies } [a, [b,c]] = [[a,b],{}^bc].&
\end{align}

@@

%To show that $f_m$ is an isomorphism we construct a map $g_m \colon G_{m+1} \to G_m$ in the opposite direction.
%We define it on generators $\mathcal{X}_{m+1}$ and then verify that it is well defined.
%Clearly, $g_m$ should be defined as the identity map on the subset of generators $\mathcal{X}_m \subseteq \mathcal{X}_{m+1}$.
%Now if $x_{\alpha}(\xi)$ is an element of $\mathcal{X}_{m+1}\setminus \mathcal{X}_m$
%there exist $\xi' \in B_m$ and $\alpha'\in \Phi$ such that $\xi = t\xi'$ and $\alpha-\alpha'\in\Phi$, 
% therefore we can set 
%\[g_m(x_\alpha(\xi)) = [x_{\alpha-\alpha'}(\xi'), x_{\alpha'}(N_{\alpha-\alpha',\alpha'} t)].\]
%It follows from~\eqref{item:i} that this definition does not depend on the choice of $\alpha'$.

Let us show that elements $S_\alpha(\xi)$ satisfy the relations $(a_{m+1})$, $(b_{m+1})$ and $(c_{m+1})$.
From \eqref{eq:H1ii} and~\eqref{item:ii} it follows that $S_\alpha(\xi)$  satisfy $(a_{m+1})$ and hence $(b_{m+1})$ in the special case $\alpha=\alpha'$.
\begin{comment}
\begin{lemma} For every $\rA_3$-triple $(\alpha, \beta, \gamma)$ in $\Phi$ and $a,b,c \in A$ the following relation holds in $G_1$:
\begin{align} 
 [x_{\alpha}(a), [x_{\beta}(b t), x_{\gamma}(c t)]] & = [x_{\beta+\alpha}(N_{\beta, \alpha}a b t), x_{\gamma}(ct)].
\end{align}
 \end{lemma}
\begin{proof}
 \begin{align*}
   [x_{\alpha}(a), [x_{\beta}(b t), x_{\gamma}(c t)]] = [[x_{\alpha}(a), x_{\beta}(b t)], {}^{x_{\beta}(b t)}\!x_{\gamma}(ct)] &  \text{ by~\eqref{eq:H1iii} and~$(b_1)$} \\ 
                                                       = {}^{x_{\beta}(b t)}[x_{\alpha +\beta}(N_{\alpha,\beta}abt), x_{\gamma}(ct) ] & \text{ by $(b_1)$ and $(c_1)$}\\ 
                                                       = [x_{\alpha +\beta}(N_{\alpha,\beta}abt), x_{\gamma}(ct)] & \text{ by $(d_1)$. \qedhere}
 \end{align*}  
\end{proof}
\end{comment}

\begin{lemma} \label{lem:bmp1} Suppose that $m\geq 1$. For every $\alpha, \alpha' \in \Phi$ such that $\alpha+\alpha' \in \Phi$ and
 $a\in A$, $\xi \in B_d$, $d \leq 0$ the following relation holds in $G_m$: 
\begin{equation} \nonumber
[x_\alpha(\xi), S_{\alpha'}(at^{m+1})] = [x_\alpha(t\xi), x_{\alpha'}(at^m)] = S_{\alpha+\alpha'}(N_{\alpha,\alpha'}a\xi t^{m+1}).
\end{equation}
\end{lemma}
\begin{proof}
By \cite[Lemma~3.1.2]{RS76} we may assume without loss of generality that $\alpha'=\beta$ for some $\rA_3$-triple $(\alpha, \beta, \gamma)$.
\begin{align*}
   [x_\alpha(\xi), S_\beta(at^{m+1})] = [x_\alpha(\xi), [x_{\beta + \gamma}(t), x_{-\gamma}(a't^m)]]
   &  \text{ by~\eqref{item:i},\eqref{eq:defmpn} for $a' = N_{\beta+\gamma, -\gamma} a$} \\ 
 = [x_{\alpha+\beta+\gamma}(\epsilon t\xi), {}^{x_{\beta+\gamma}(t)}x_{-\gamma}(a't^m)]             
 &  \text{ by~\eqref{eq:H1iii}, for $\epsilon=N_{\alpha, \beta+\gamma}$} \\
 = {}^{x_{\beta+\gamma}(t)}[x_{\alpha+\beta+\gamma}(\epsilon t\xi), x_{-\gamma}(a't^m)]             
 &  \text{ by $(b_1)$}
\end{align*} 
Denote by $R$ the expression in the right hand side of the above formula.
\begin{enumerate}
 \item \label{case:cm-1} Case $m=1$, $d = 0$.
 \begin{align*}
   R  = [x_{\alpha +\beta + \gamma}(\epsilon t\xi), x_{-\gamma}(a't)] 
   & \text{ by $(d_1)$}\\
      = {}^{x_{\beta+\gamma}(1)}[x_{\alpha +\beta + \gamma}(\epsilon t\xi), x_{-\gamma}(a't)] 
   & \text{ by~\eqref{eq:H1ii}, $(b_1)$, $(c_1)$} \\
      = [x_\alpha(\xi), [x_{\beta + \gamma}(1), x_{-\gamma}(a't^m)]] 
   & \text{ by $(b_1)$, \eqref{eq:H1iii}}\\
      = [x_\alpha(t\xi), x_\beta(at)]
   & \text{ by~\eqref{item:i},\eqref{eq:defmpn}.}
 \end{align*}  
 %This uses $N_{\alpha, \beta+\gamma} N_{\alpha+\beta+\gamma, -\gamma} N_{\beta+\gamma, -\gamma} = N_{\alpha, \beta}$. 
 %Why is this true?

 \item \label{case:cm-2} Case $m\geq 2$ or $m=1$, $d < 0$.
 \begin{align*}
 R = {}^{x_{\beta+\gamma}(t)}[x_{\alpha+\beta+\gamma}(\epsilon t^2\xi), x_{-\gamma}(a't^{m-1})]     
 &  \text{ by~\eqref{item:i} if $k=0$ or~\eqref{eq:cm} if $k <0$} \\
 = [[x_\alpha(t\xi), x_{\beta+\gamma}(t)], {}^{x_{\beta+\gamma}(t)} x_{-\gamma}(a't^{m-1})]         
 &  \text{ by~$(b_2)$, $(c_2)$ or by~\eqref{item:ii},\eqref{eq:defmpn} if $m=1$} \\
 = [x_\alpha(t\xi), x_\beta(at^m)]                                                                  
 &  \text{ by~\eqref{eq:H1iii}. \qedhere}
\end{align*}

@@

Now suppose  $0 \leq d \leq m+1$. Clearly \eqref{lem:bmp1} and \eqref{eq:defmpn} (in the case $1\leq d\leq m$) or~\cref{lem:bmp1} (in the case $d=0,m+1$) imply that
\begin{equation} \label{eq:verify-bmp1'} 
[S_\alpha(\xi), S_{\alpha'}(\eta)] = S_{\alpha+\alpha'}(N_{\alpha,\alpha'}\xi \eta),\ \xi \in B_d, \eta \in B_{m+1-d}.
\end{equation}

@@

\begin{lemma}
Elements $S_\alpha(\xi)$ satisfy relations $(b_{m+1})$.
 \end{lemma}
\begin{proof}
We need to show that $S_\alpha(\xi)$ satisfy $\angle(d, m+1)$ and $\bot(d, m+1)$ for $d\leq m+1$.
\begin{enumerate}
 \item Relations $\angle(d, m+1)$, $d\leq m$ follow from~\cref{claim1}. 
 \item Relations $\bot(d, m+1)$ with $d\leq 0$ can verified by direct computation:
 \begin{align*} 
 [x_\alpha(\xi), S_{\gamma}(bt^{m+1})] = [x_\alpha(\xi), [x_{\beta+\gamma}(b't^m), x_{-\beta}(t)]] & \text{ by~\eqref{item:i},\eqref{eq:defmpn}} \\
 = [[x_\alpha(\xi), x_{\beta+\gamma}(b't^m)], {}^{x_{\beta+\gamma}(b't^m)}\!x_{-\beta}(t)] &
 \text{ by \eqref{eq:H1iii} and $(b_1)$} \\
 = {}^{x_{\beta+\gamma}(b't^m)}\![x_{\alpha+\beta+\gamma}(b't^m\xi), x_{-\beta}(t)] &
 \text{ by $(b_m)$ and $(c_m)$} \\
 = 1 & \text{ by $(b_m)$.}
\end{align*}
 \item Relations $\bot_0(d, m+1)$ with $1\leq d\leq m$ follow from~\cref{claim2}.
 \item Relations $\angle(m+1, m+1)$ follow from~\cref{claim1}, consequently relations $\bot_0(m+1, m+1)$ follow from~\cref{claim2}.
 \item Relations $\bot(d, m+1)$, $0\leq d\leq m+1$ follow from~\cref{claim3} and the previous two assertions.
\end{enumerate}
\end{proof}

@@

Notice that the second assertion follows from the first one.
Indeed, this would give the required isomorphism $G_1 \cong \St(\Phi, B)$ since $\langle \XX{\infty} \mid \RR{\infty} \rangle$ is evidently isomorphic to $\St(\Phi, B)$.
%Denote by $F(S)$ the free group on the generating set $S$.
%It is clear that $\RR{n}\subseteq\RR{n+1}$, moreover $\langle \RR{n} \rangle = \langle \RR{\infty}\rangle \cap F(\XX{n})$.
%well, not so clear, actually :)

\end{proof}

>>>>

\documentclass[oneside, 10pt]{amsart}
\usepackage{amscd, amsmath, amssymb, amsthm, amsfonts, amstext, verbatim, mathtools, xfrac, microtype, nameref, thmtools}
\usepackage[breaklinks=true]{hyperref}
\usepackage[hyperref=true, backend=bibtex, firstinits=true, citestyle=numeric-comp, sortlocale=en_US, url=false, doi=false, eprint=true, maxbibnames=4]{biblatex}
\usepackage[capitalize]{cleveref}
\usepackage[matrix,arrow,curve]{xy}
\usepackage{tikz}
\usepackage{enumitem}

\addbibresource{paper.bib}
\renewbibmacro*{volume+number+eid}{\ifentrytype{article}{\- \iffieldundef{volume}{}{Vol.~\printfield{volume},}\iffieldundef{number}{}{ No.~\printfield{number},}}}
\renewbibmacro{in:}{\ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}}
\newbibmacro{string+doi}[1]{\iffieldundef{doi}{\iffieldundef{url}{#1}{\href{\thefield{url}}{#1}}}{\href{http://dx.doi.org/\thefield{doi}}{#1}}}
\DeclareFieldFormat[article, inproceedings, inbook, book, thesis]{title}{\usebibmacro{string+doi}{\mkbibquote{#1}}}
\renewcommand*{\bibfont}{\footnotesize}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\T}{T}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\colim}{colim}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\SL}{SL}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\rk}{rk}
\newcommand{\catname}[1]{{\normalfont\textbf{#1}}}

@@

We need one additional technical definition. We call an ordered triple of roots $(\alpha, \beta, \gamma) \in \Phi^3$ an {\it $\rA_3$-triple}
 if these vectors form a basis of a root subsystem $\Psi \subseteq \Phi$ of type $\rA_3$, and moreover if the Dynkin diagram of $\Psi$ with respect to this basis has the form
\begin{tikzpicture}
\draw[fill=black] 
(0,0) circle [radius=.1] node [below] {$\alpha$} --
(1,0) circle [radius=.1] node [below] {$\beta$} --
(2,0) circle [radius=.1] node [below] {$\gamma$}
\end{tikzpicture}.

@@

We will use the following commutator identities (cf.~\cite[H1]{Re75}):
\begin{align}
 \label{eq:H1ii}  [ab, c] = {}^a[b, c] \cdot [a,c];&\\ %= [a,[b,c]]\cdot [b,c] \cdot [a,c];&\\
 \label{eq:H1iii} [a,c]   = 1    \text{ implies } [a, [b,c]] = [[a,b],{}^bc].&
\end{align}

@@

Let us show that elements $S_\alpha(\xi)$ satisfy the relations $(a_{m+1})$, $(b_{m+1})$ and $(c_{m+1})$.
From \eqref{eq:H1ii} and~\eqref{item:ii} it follows that $S_\alpha(\xi)$ satisfy $(a_{m+1})$ and hence $(b_{m+1})$ in the special case $\alpha=\alpha'$.
\begin{lemma} \label{lem:bmp1} Suppose that $m\geq 1$. For every $\alpha, \alpha' \in \Phi$ such that $\alpha+\alpha' \in \Phi$ and
 $a\in A$, $\xi \in B_d$, $d \leq 0$ the following relation holds in $G_m$: 
\begin{equation} \nonumber
[x_\alpha(\xi), S_{\alpha'}(at^{m+1})] = [x_\alpha(t\xi), x_{\alpha'}(at^m)] = S_{\alpha+\alpha'}(N_{\alpha,\alpha'}a\xi t^{m+1}).
\end{equation}
\end{lemma}
\begin{proof}
By \cite[Lemma~3.1.2]{RS76} we may assume without loss of generality that $\alpha'=\beta$ for some $\rA_3$-triple $(\alpha, \beta, \gamma)$.
\begin{align*}
   [x_\alpha(\xi), S_\beta(at^{m+1})] = [x_\alpha(\xi), [x_{\beta + \gamma}(t), x_{-\gamma}(a't^m)]]
   &  \text{ by~\eqref{item:i},\eqref{eq:defmpn} for $a' = N_{\beta+\gamma, -\gamma} a$} \\ 
 = [x_{\alpha+\beta+\gamma}(\epsilon t\xi), {}^{x_{\beta+\gamma}(t)}x_{-\gamma}(a't^m)]             
 &  \text{ by~\eqref{eq:H1iii}, for $\epsilon=N_{\alpha, \beta+\gamma}$} \\
 = {}^{x_{\beta+\gamma}(t)}[x_{\alpha+\beta+\gamma}(\epsilon t\xi), x_{-\gamma}(a't^m)]             
 &  \text{ by $(b_1)$}
\end{align*} 
Denote by $R$ the expression in the right hand side of the above formula.
\begin{enumerate}
 \item \label{case:cm-1} Case $m=1$, $d = 0$.
 \begin{align*}
   R  = [x_{\alpha +\beta + \gamma}(\epsilon t\xi), x_{-\gamma}(a't)] 
   & \text{ by $(d_1)$}\\
      = {}^{x_{\beta+\gamma}(1)}[x_{\alpha +\beta + \gamma}(\epsilon t\xi), x_{-\gamma}(a't)] 
   & \text{ by~\eqref{eq:H1ii}, $(b_1)$, $(c_1)$} \\
      = [x_\alpha(\xi), [x_{\beta + \gamma}(1), x_{-\gamma}(a't^m)]] 
   & \text{ by $(b_1)$, \eqref{eq:H1iii}}\\
      = [x_\alpha(t\xi), x_\beta(at)]
   & \text{ by~\eqref{item:i},\eqref{eq:defmpn}.}
 \end{align*}  
 \item \label{case:cm-2} Case $m\geq 2$ or $m=1$, $d < 0$.
 \begin{align*}
 R = {}^{x_{\beta+\gamma}(t)}[x_{\alpha+\beta+\gamma}(\epsilon t^2\xi), x_{-\gamma}(a't^{m-1})]     
 &  \text{ by~\eqref{item:i} if $k=0$ or~\eqref{eq:cm} if $k <0$} \\
 = [[x_\alpha(t\xi), x_{\beta+\gamma}(t)], {}^{x_{\beta+\gamma}(t)} x_{-\gamma}(a't^{m-1})]         
 &  \text{ by~$(b_2)$, $(c_2)$ or by~\eqref{item:ii},\eqref{eq:defmpn} if $m=1$} \\
 = [x_\alpha(t\xi), x_\beta(at^m)]                                                                  
 &  \text{ by~\eqref{eq:H1iii}. \qedhere}
\end{align*}

@@

Now suppose  $0 \leq d \leq m+1$. Clearly \eqref{item:i} and \eqref{eq:defmpn} (in the case $1\leq d\leq m$) or~\cref{lem:bmp1} (in the case $d=0,m+1$) imply that
\begin{equation} \label{eq:verify-bmp1'} 
[S_\alpha(\xi), S_{\alpha'}(\eta)] = S_{\alpha+\alpha'}(N_{\alpha,\alpha'}\xi \eta),\ \xi \in B_d, \eta \in B_{m+1-d}.
\end{equation}

@@

\begin{lemma}
Elements $S_\alpha(\xi)$ satisfy relations $(b_{m+1})$.
 \end{lemma}
\begin{proof}
We must show that $S_\alpha(\xi)$ satisfy $\angle(d, m+1)$ and $\bot(d, m+1)$ for $d\leq m+1$.
\begin{enumerate}
 \item Relations $\angle(d, m+1)$, $d\leq m$ follow from~\cref{claim1}. 
 \item Relations $\bot(d, m+1)$ with $d\leq 0$ can verified by direct computation:
 \begin{align*} 
 [x_\alpha(\xi), S_{\gamma}(bt^{m+1})] = [x_\alpha(\xi), [x_{\beta+\gamma}(b't^m), x_{-\beta}(t)]] & \text{ by~\eqref{item:i},\eqref{eq:defmpn}} \\
 = [[x_\alpha(\xi), x_{\beta+\gamma}(b't^m)], {}^{x_{\beta+\gamma}(b't^m)}\!x_{-\beta}(t)] &
 \text{ by \eqref{eq:H1iii} and $(b_1)$} \\
 = {}^{x_{\beta+\gamma}(b't^m)}\![x_{\alpha+\beta+\gamma}(b't^m\xi), x_{-\beta}(t)] &
 \text{ by $(b_m)$ and $(c_m)$} \\
 = 1 & \text{ by $(b_m)$.}
\end{align*}
 \item Relations $\bot_0(d, m+1)$ with $1\leq d\leq m$ follow from~\cref{claim2}.
 \item Relations $\angle(m+1, m+1)$ follow from~\cref{claim1}, consequently relations $\bot_0(m+1, m+1)$ follow from~\cref{claim2}.
 \item Relations $\bot(d, m+1)$, $0\leq d\leq m+1$ follow from~\cref{claim3} and the previous two assertions.
\end{enumerate}
\end{proof}

@@

Denote by $G_\infty$ the group presented by generators $\mathcal{X}_\infty$ and relations $(a_\infty)$, $(b_\infty)$, $(c_\infty)$.
By the above argument there is an isomorphism between $G_1$ and $\colim\limits_{n\to\infty} G_n \cong G_\infty$.

Denote by $i\colon G_\infty \to \St(\Phi, B)$ the map induced by the obvious embedding of $\mathcal{X}_\infty$ into the set of 
 Steinberg generators of $\St(\Phi, B)$.  
Using~\eqref{eq:H1ii} multiple times it is not hard to show that the map $j\colon \St(\Phi, B) \to G_\infty$ given by
 $j(x_\alpha(b)) = \prod x_\alpha(b_i)$, where $b = \sum b_i$ is the decomposition of $b$ into a finite sum of its homogeneous components $b_i\in B_i$, is a well-defined map.
It is clear that $i$ and $j$ are mutually inverse. 
\end{proof}
